---
title: "Psychonomics 2023: ALL SHAPES: OLD VS NEW"
author: "Joanna Morris"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's import our data

```{r}
library(readr)
subjectProfiles <- read_csv("M21_subjectProfiles_hampshire_2.csv")
n400 <- read_csv("m21_vsl_300500_150050.csv")
```

Next lets join the subject reading profile data and erp data into a single dataframes

```{r}
library(dplyr)
library(tidyr)
n400_subj<-left_join(n400, subjectProfiles, by = "SubjID")
```


Now let's divide this data frame into four separate ones—for responses to 1st, 2nd, 3rd and all shapes
```{r}
library(tidyr)
n400_subj <- separate_wider_delim(n400_subj,
                                  binlabel, 
                                  delim = "_", 
                                  names = c("TrialType", "file"), 
                                  too_few = "align_start")
n400_first <- filter(n400_subj, file == 1)
n400_second <- filter(n400_subj, file == 2)
n400_third <- filter(n400_subj, file == 3)
n400_all <- filter(n400_subj, is.na(n400_subj$file))
n400_all$file <- NULL
```


Now for the `n400_all` dataframe which is what we will be analysing, we need to  extract just the bins and channels that we intend to analyse.  

```{r}
channels <-  c(3, 2, 25, 7, 20, 21, 12, 11, 16)
n400_subset <- n400_all |> filter(chindex %in% channels) 
n400_subset_a <- separate_wider_position(n400_subset,
                                       cols = chlabel, 
                                       widths = c(anteriority = 1, laterality= 1),
                                       cols_remove = FALSE) # separate by dividing into two columns

n400_subset_b <- mutate(n400_subset_a, 
                        Laterality = case_when((laterality == "z"|laterality == "Z") ~ "Midline",
                                             laterality == "3" ~ "Left", 
                                             laterality == "4" ~ "Right"),
                        Anteriority = case_when(anteriority == "F" ~ "Frontal",
                                                anteriority == "C" ~ "Central",
                                                anteriority == "P" ~ "Parietal"),
                        .keep = "unused")

n400_subset_oldnew <- n400_subset_b |> filter(TrialType == "New" | TrialType == "Old")
```

Now we can run our anova. First let's check to make sure we have no empty cells

```{r}
(responses_table <- table(n400_subset_oldnew$ReadingProfile, 
                          n400_subset_oldnew$Anteriority,
                          n400_subset_oldnew$Laterality))
```

Then we run the anova
```{r}

library(ez)
ezANOVA(n400_subset_oldnew,
        dv = value,
        wid = SubjID,
        within = .(TrialType,Anteriority,Laterality),
        between = ReadingProfile,
        type = 3)
```

We have a significant ReadingProfile x Trial Type by Laterality Interaction.  Let's examine the means for the conditions.

```{r}
library(knitr)
library(plotrix)
kable(means_table <- n400_subset_oldnew |> 
  group_by(Laterality, ReadingProfile, TrialType ) |>
  summarise(Mean = mean(value), 
            SE = std.error(value), 
            SD = sd(value), 
            Max = max(value), 
            Min = min(value)), digits = 2)

```

Now, let's run some post-hoc tests
```{r}
library(pander)
pander(t.test(value ~ TrialType
       , filter(n400_subset_oldnew, ReadingProfile == "SEMANTIC" & Laterality == "Left" )
       , paired=TRUE))

library(pander)
pander(t.test(value ~ TrialType
       , filter(n400_subset_oldnew, ReadingProfile == "SEMANTIC" & Laterality == "Right" )
       , paired=TRUE))

library(pander)
pander(t.test(value ~ TrialType
       , filter(n400_subset_oldnew, ReadingProfile == "ORTHOGRAPHIC" & Laterality == "Left" )
       , paired=TRUE))

library(pander)
pander(t.test(value ~ TrialType
       , filter(n400_subset_oldnew, ReadingProfile == "ORTHOGRAPHIC" & Laterality == "Right" )
       , paired=TRUE))

```

Let's plot the means
```{r fig.dim = c(8, 8)}
library(ggplot2)
library(ggsci)
n400_right <- filter(n400_subset_oldnew, Laterality == "Right") |> 
  group_by(ReadingProfile, TrialType) |>
  summarise(mean = mean(value), se = std.error(value))|> 
  mutate(lwr = mean - (1.96*se), upr = mean + (1.96*se))


n400_left <- filter(n400_subset_oldnew, Laterality == "Left") |> 
  group_by(ReadingProfile, TrialType) |>
  summarise(mean = mean(value), se = std.error(value))|> 
  mutate(lwr = mean - (1.96*se), upr = mean + (1.96*se))


plot1<- ggplot(n400_left, aes(x = ReadingProfile, 
                         y = mean,
                         ymin = lwr,
                         ymax = upr,
                         fill = TrialType, 
                         group = TrialType)) + 
  geom_bar(position=position_dodge(), stat="identity", colour = "black") + 
  geom_errorbar(width = .2, position = position_dodge(.9))  +
  scale_fill_npg() + ylab("Voltage in microvolts") +
  ggtitle("Reading Profile by Trial Type Interaction: Left") + theme_classic()

plot2 <- ggplot(n400_right, aes(x = ReadingProfile, 
                         y = mean,
                         ymin = lwr,
                         ymax = upr,
                         fill = TrialType, 
                         group = TrialType)) + 
  geom_bar(position=position_dodge(), stat="identity", colour = "black") + 
  geom_errorbar(width = .2, position = position_dodge(.9))  +
  scale_fill_npg() + ylab("Voltage in microvolts") +
  ggtitle("Reading Profile by Trial Type Interaction: Right") + theme_classic()
 

library(gridExtra)
grid.arrange(plot1, plot2, ncol=1)
```

